Measurement and Probability in Fuzzy Quantum Theories

### 2017-02-20 prop_iop.tex
more on intro

Textbook quantum mechanics asserts that, given a quantum spin system
described by the wavefunction
$\ket{\psi}=\frac{\sqrt{\pi}}{2}\fket{\ps\frac{1}{2}}+\frac{\sqrt{4-\pi}}{2}\fket{\minus\frac{1}{2}}$
and the observable $S_z = \frac{\hbar}{2}(\fproj{\ps\frac{1}{2}}-\fproj{\minus\frac{1}{2}})$,
the probability of observing the eigenvalue $\ps\frac{\hbar}{2}$ (with
post-measurement state $\fket{\ps\frac{1}{2}}$) is $\frac{\pi}{4}$ and
the probability of observing $\minus\frac{\hbar}{2}$ (with post-measurement
state $\fket{\minus\frac{1}{2}}$) is $1-\frac{\pi}{4}$. According to the
frequency interpretation of probability, the above assertion implies
that when performing a very large number of experiments on quantum
systems described by wavefunction $\ket{\psi}$, a fraction
$\frac{\pi}{4}$ of the experiments will result in the observation
$\ps\frac{\hbar}{2}$ and the remaining fraction $1-\frac{\pi}{4}$ will
result in the observation $\minus\frac{\hbar}{2}$.

In reality, i.e., in an actual experiment, simulation, or computation,
bounded by space, time, and energy resources, the situation is much
fuzzier: (i) we cannot \emph{exactly} prepare the state $\psi$; (ii)
we cannot build a measurement apparatus corresponding \emph{exactly}
to observable $S_z$; and (iii) we cannot prepare \emph{exact} replicas
of any state. In other words, we can only perform a limited number of
experiments on quantum systems closely related to wavefunction
$\ket{\psi}$ and using an apparatus that closely approximates
observable $S_z$. Given any fixed resources, these experiments will
\emph{not} result in the \emph{exact} fraction $\frac{\pi}{4}$ of
observations $\ps\frac{\hbar}{2}$ and $1-\frac{\pi}{4}$ of
observations $\minus\frac{\hbar}{2}$. Indeed, the state of the art
algorithms for computing the $n$th binary digit of~$\pi$ require on
the order of $O\left(n\log^{O\left(1\right)}\left(n\right)\right)$
operations~\cite{journals/moc/BaileyBP97} which means that simply
referring to the $n$th digit of $\pi$ requires more and more resources
as~$n$ gets larger.

All of this is well-known and holds in both the classical and quantum
world. In the classical world, these observations are believed to be
well-understood approximations to an idealized reality. The question
in the quantum world is more subtle and it is not evident at all that
it does not alter the very foundations of quantum mechanics. Indeed
taking such resource bounds into consideration is what founded
computer science as a discipline and has become crucial for
understanding the very nature of computation. Following
Feynman~\cite{Feynman1982Simulating}, Landauer~\cite{Landauer1996188},
and others, one might argue that this resource-bounded perspective is
also crucial for understanding the very nature of physical processes.

In previous work, \ldots

>>>>

Textbook quantum mechanics asserts that, given a quantum spin system
described by the wavefunction
$\ket{\psi}=\frac{\sqrt{\pi}}{2}\fket{\ps\frac{1}{2}}+\frac{\sqrt{4-\pi}}{2}\fket{\minus\frac{1}{2}}$
and given the observable
$S_z =
\frac{\hbar}{2}(\fproj{\ps\frac{1}{2}}-\fproj{\minus\frac{1}{2}})$,
then the probability of observing the eigenvalue $\ps\frac{\hbar}{2}$
(with post-measurement state $\fket{\ps\frac{1}{2}}$) is
$\frac{\pi}{4}$ and the probability of observing the eigenvalue
$\minus\frac{\hbar}{2}$ (with post-measurement state
$\fket{\minus\frac{1}{2}}$) is $1-\frac{\pi}{4}$. According to the
frequency interpretation of probability, the above assertion implies
that when performing a very large number of experiments on quantum
systems described by wavefunction $\ket{\psi}$, a fraction
$\frac{\pi}{4}$ of the experiments will result in the observation
$\ps\frac{\hbar}{2}$ and the remaining fraction $1-\frac{\pi}{4}$ will
result in the observation $\minus\frac{\hbar}{2}$.

In reality, i.e., in an actual experiment, simulation, or computation,
bounded by space, time, and energy resources, the situation is much
fuzzier: (i) we cannot \emph{exactly} prepare the state $\ket{\psi}$;
(ii) we cannot build a measurement apparatus corresponding
\emph{exactly} to observable $S_z$; and (iii) we cannot prepare
\emph{exact} replicas of any state. In other words, we can only
perform a limited number of experiments on quantum systems closely
related to wavefunction $\ket{\psi}$ and using an apparatus that
closely approximates observable $S_z$. Given any fixed resources,
these experiments will \emph{not} result in the \emph{exact} fraction
$\frac{\pi}{4}$ of observations $\ps\frac{\hbar}{2}$ and
$1-\frac{\pi}{4}$ of observations $\minus\frac{\hbar}{2}$. In fact,
these experiments can be viewed as a method (algorithm) to calculate
the value $\pi$ and it is known that the state of the art algorithms
for computing the $n$th binary digit of~$\pi$ require on the order of
$O\left(n\log^{O\left(1\right)}\left(n\right)\right)$
operations~\cite{journals/moc/BaileyBP97}. In other words, calculating
the $n$th digit of $\pi$ appears to require more and more physical
resources as~$n$ gets larger.

All of this is well-known and holds in both the classical and quantum
worlds. In the classical world, these observations are believed to be
well-understood approximations to an idealized independent reality
that can continuously approached with more and more resources. The
question in the quantum world is more subtle; it is not evident at all
that this resource-aware perspective does not alter the very
foundations of quantum mechanics. Indeed taking such resource bounds
into consideration is what founded computer science as a discipline
and has become crucial for understanding the very nature of
computation. Following Feynman~\cite{Feynman1982Simulating},
Landauer~\cite{Landauer1996188}, and others, one might argue that this
resource-bounded perspective is also crucial for understanding the
very nature of physical processes.

In previous work, we analyzed the wavefunction description of quantum
systems from a resource-bounded perspective. Technically, instead of
postulating that quantum states are rays in a Hilbert space, i.e., we
instead used special finite fields whose size reflected the available
resources. \ldots
